|

### 2021-12-16 front-and-back-matter/99-appendices.Rmd
Other edits

@@

A copy of the risk of bias tool used to assess Mendelian randomisation studies is presented in Table \@ref(mrTool-table), below.

@@

<!----------------------------------------------------------------------------->
(ref:prismaTab-caption) __PRISMA Checklist__ - Following best practice guidelines, the Section/Table/Figure number were important details on the systematic review can be found are listed.

@@

@@

(ref:drDemeniaTC-cap) __Dose-response meta-analysis of total cholesterol on all-cause dementia__

@@

(ref:drAdLDL-cap) __Dose-response meta-analysis of low-density lipoprotein cholesterol on Alzheimer's disease__

@@

Unfortunately, given the nature of the data underpinning the CPRD (Chapter \@ref(cprd-analysis-heading)) and IPD (Chapter \@ref(ipd-heading)) analyses, it is not possible to make it publicly available. Access to the CPRD data is dependent on an application to its managing body, while acces to the cohorts within the DPUK requires approval from both the DPUK and inidividual cohort owners. In the absence of data sharing, the best option to enable review and increase transparency is the open sharing of the code used to perform the analysis.[@goldacre2019]

@@

OVer the course of my PhD, I kept a record of attempted

Short section detailing the things I tried to do but which did not work:

- Failed to properly scope the literature and correctly estimate the number of studies returned by the search

- Failed IV analysis of anti-HC drugs because of low numbers of non-statins.

- Failed to realise that the initial code list used for the positive control outcome of IHD in the CPRD analysis (Chapter \@ref(cprd-analysis-heading))

- Unsuccessful grant application with Andy

- Unsuccessful grant application with Chris Penfold

- Unsuccessful grant application with Matt

- Instrumental variable analysis of lipid regulation use using physicians prescribing preference.

- 2 x unsuccessful conference abstracts

<!-- TODO Include the early analysis we did here, that did not work as a result of limited scope. Could include some of the genetic work also. Will additionally need to put the relevant materials from the paper in the  -->

@@

## Producing risk-of-bias visualisations with `robvis` {#appendix-robvis}

<!-- TODO Need to complete this sections -->
The `robvis` package provides functions to convert a risk-of-bias assessment summary table into a summary plot or a traffic-light plot, formatted based on the specific risk-of-bias assessment tool used.

### Templates

`robvis` currently contains templates for the following tools:

* ROB2
* ROBINS-I
* QUADAS-2
* ROB1
* General

### Data expectations

For the direction bias plots, the data expected by the tools is more comprehensive.

@@

<!----------------------------------------------------------------------------->
(ref:biasDirectionData-caption) __Example bias direction dataset__ - Example risk-of-bias data illustrating the structure required for bias direction plots. $d_ij$, $d_id$, and $d_it$ define the extent (Serious, Moderate, Low), direction (right or left), and type (additive or proportional) of bias in the $i$th domain. Note that, because of the wide nature of the data, I have abbreviated the data in order to display it. A key for interpreting the data can be found below the table.

@@

### Simple visualisations

```{r robAppendixFigures, include = FALSE}
```

```{r, eval = FALSE}
rob_summary(data = data_rob2, tool = "ROB2")
```

<!----------------------------------------------------------------------------->
(ref:robSummary-cap) __Example risk-of-bias summary barplot__ - Illustration of a simple summary barplot created using the `data_rob` example risk-of-bias dataset shown in Table \@ef(tab:biasData-table) above. Note that this plot is unweighted and shows the _proportion of studies_ at each level of bias, but an option exists with `robvis` to create a barplot weighted by the _proportion of evidence_ at each level of bias.

(ref:robSummary-scap) Example risk-of-bias summary barplot

```{r robSummary, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:robSummary-cap)', out.width='80%', fig.scap='(ref:robSummary-scap)'}
knitr::include_graphics(file.path("figures/appendix/robSummary.png"))
```
<!----------------------------------------------------------------------------->

```{r, eval = FALSE}
rob_traffic_light(data = data_rob2,
                  tool = "ROB2",
                  psize = 10)
```

<!----------------------------------------------------------------------------->
(ref:robTraffic-cap) Example risk-of-bias traffic-light plot - Illustration of a simple traffic light plot created using the `data_rob` example risk-of-bias dataset shown in Table \@ef(tab:biasData-table) above.

(ref:robTraffic-scap) Example risk-of-bias traffic-light plot

```{r robTraffic, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:robTraffic-cap)', out.width='80%', fig.scap='(ref:robTraffic-scap)'}
knitr::include_graphics(file.path("figures/appendix/robTraffic.png"))
```
<!----------------------------------------------------------------------------->

@@

<!-- TODO May need to reiterate contributions here -->

@@

\includepdf[pages=-,link=true,pagecommand={\thispagestyle{empty}}]{published/cprd_medrxiv.pdf}
\includepdf[pages=-,pagecommand={\thispagestyle{empty}}]{published/medrxivr.pdf}
\includepdf[pages=-,pagecommand={\thispagestyle{empty}}]{published/robvis.pdf}
\includepdf[pages=-,pagecommand={\thispagestyle{empty}}]{published/DAS_comparison.pdf}
\includepdf[pages=-,pagecommand={\thispagestyle{empty}}]{published/prevsci.pdf}

>>>>

A copy of the risk of bias tool used to assess Mendelian randomisation studies is presented in Table \@ref(tab:mrTool-table), below.

@@

@@

<!----------------------------------------------------------------------------->
(ref:prismaTab-caption) __PRISMA Checklist__ - Following best practice guidelines, the Section/Table/Figure number where important details on the systematic review can be found are listed.

@@

### Funnel plots {#appendix-funnel-plots}

```{r funnelPlots, include = FALSE}
```

<!----------------------------------------------------------------------------->
(ref:funnerlPlots-cap) __Funnel plots__ - To assess the potential for small study effects, funnel plots were produced for the three meta-analyses reported in Chapter \@ref(sys-rev-results-heading) that contained more than 10 results. Visual inspection of the plots did not indicate the presence of small study effect, which was supported in each case by the result of Egger's regression test for funnel plot asymmetry.

(ref:funnerlPlots-scap) Funnel plots

```{r funnerlPlots, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:funnerlPlots-cap)', out.width='100%', fig.scap='(ref:funnerlPlots-scap)'}
knitr::include_graphics(file.path("figures/appendix/funnel_composite.pdf"))
```
<!----------------------------------------------------------------------------->

@@

(ref:drDemeniaTC-cap) __Dose-response meta-analysis of total cholesterol on all-cause dementia__ - Non-randomised studies examining the association of total cholesterol with all-cause dementia across several categories or "doses" of exposure were synthesised using a dose response meta-analysis. The reference dose was defined _a priori_ as 200 _mg/dL_, the cut-off of the "Normal"/"Optimal" category for total cholesterol as detailed in Table \@ref(tab:lipidLevels-table).

@@

(ref:drAdLDL-cap) __Dose-response meta-analysis of low-density lipoprotein cholesterol on Alzheimer's disease__ - Non-randomised studies examining the association of LDL with Alzheimer's disease across several categories or "doses" of exposure were synthesised using a dose response meta-analysis. The reference dose was defined _a priori_ as 100 _mg/dL_, the cut-off of the "Normal"/"Optimal" category for LDL as detailed in Table \@ref(tab:lipidLevels-table).

@@

Unfortunately, given the nature of the data underpinning the CPRD (Chapter \@ref(cprd-analysis-heading)) and IPD (Chapter \@ref(ipd-heading)) analyses, it is not possible to make it publicly available. Access to the CPRD data is dependent on an application to its managing body, while access to the cohorts within the DPUK requires approval from both the DPUK and individual cohort owners. In the absence of data sharing, the best option to enable review and increase transparency is the open sharing of the code used to perform the analysis.[@goldacre2019]

@@

As part of a reflective process, during the course of my PhD I kept a record of things (analytical mishaps, failed experiments, and unsuccessful grant applications) I attempted but which did not work. These are summarised briefly here:

- Prior to the analysis performed in Chapter \@ref(cprd-analysis-heading), I attempted an analysis using physicians prescribing preference as an instrumental variable to compare different lipid-regulating agent classes for dementia prevention. However, this analysis was not pursued due to the very small number of patients prescribed a lipid-regulating agent other than statins.

- I initially failed to realise that a code list I was using for the positive control outcome of ischemic heart disease (IHD) in the CPRD analysis (Chapter \@ref(cprd-analysis-heading)) included only very severe IHD outcomes. This resulted in the positive protective control analysis giving substantially inflated hazard ratios (HR > 4) and led to substantial confusion in the study team.

- For the IPD analysis in Chapter \@ref(ipd-heading), I had to be reminded that (a) BMI can be calculated from weight and height, and (b) that whether you ask at 45 or 50, the age at which someone left school will be the same.

- I was involved in unsuccessful grant applications to the NIHR (with Andrew Beswick), to the Jean Golding Institute (with James Byrne and Chris Penfold) and to the PGR Hub (with Matthew Lee).

@@

## `robvis`: producing risk-of-bias visualisations {#appendix-robvis}

<!-- TODO Need to complete this sections -->
The `robvis` package provides functions to visualise risk-of-bias assessments, with the resulting plots formatted based on the specific risk-of-bias assessment tool used. The tool is documented extensively online at [https://mcguinlu.github.io/robvis/](https://mcguinlu.github.io/robvis/). As such, it is not described in detail here, other than to illustrate the data structures required for the figures presented in this thesis and to describe its reception in the evidence synthesis community.

### Data setup

For the paired forest plots presented in Chapter \@ref(sys-rev-results-heading), the tool expects a "Study" column, followed by one column per domain of bias and one column for the overall risk of bias judgement. To illustrate this, the `data_rob2` dataset which is available via the `robvis` package and contains example risk-of-bias assessments performed using the RoB2.0 tool for randomized controlled trials, is shown in Table \@ref(tab:biasData-table).

For the bias direction plots illustrated in Chapter \@ref(tri-heading), additional details on each domain of bias are required by the tool, namely the direction and type of bias in each domain. An example of the data required by `robvis` in order to produce the bias direction plots is shown in \@ref(tab:biasDirectionData-table).

```{r, echo = FALSE, warning=FALSE}

robstats <- cranlogs::cran_downloads("robvis", from = "2019-11-22", to = "2021-12-01")

robvis_downloads <- plyr::round_any(sum(robstats$count), 100, f = floor)

date <- format(as.Date("2021-12-01"),"%B %Y")
```

### Reception

The tool has been well received by the systematic review community. As of `r date`, the `robvis` R package has been downloaded more than `r medrxivr_downloads` times and the paper introducing the tool had been cited more than 450 times. A copy of this publication is available in Appendix \@ref(published-papers). Development of the tool is ongoing, as evidenced by the new bias direction plot functionality, and there are now a number of contributors to the package.

@@

<!----------------------------------------------------------------------------->
(ref:biasDirectionData-caption) __Example bias direction dataset__ - Example risk-of-bias data illustrating the structure required for bias direction plots. $d_ij$, $d_id$, and $d_it$ define the extent (Serious, Moderate, Low), direction (right or left), and type (additive or proportional) of bias in the $i$th domain, respectively Note that, because of the wide nature of the data, I have abbreviated the data in order to display it. A key for interpreting the data can be found below the table.

@@

@@

@@

<!-- \includepdf[pages=-,link=true,pagecommand={\thispagestyle{empty}}]{published/cprd_medrxiv.pdf} -->
<!-- \includepdf[pages=-,pagecommand={\thispagestyle{empty}}]{published/medrxivr.pdf} -->
<!-- \includepdf[pages=-,pagecommand={\thispagestyle{empty}}]{published/robvis.pdf} -->
<!-- \includepdf[pages=-,pagecommand={\thispagestyle{empty}}]{published/DAS_comparison.pdf} -->
<!-- \includepdf[pages=-,pagecommand={\thispagestyle{empty}}]{published/prevsci.pdf} -->

### 2021-12-16 front-and-back-matter/_abstract.Rmd
Other edits

__Background:__ Identification of causal relationships between modifiable targets and risk of dementia is central to the development of evidence-based prevention strategies. Blood lipid levels have been implicated in the aetiology of dementia by genetic linkage and functional cell biology studies, but current epidemiological evidence has yet to reach a consensus on their role in dementia risk. This thesis sought to triangulate the causal effects of lipid levels (total cholesterol, LDL-c, HDL-c, and triglycerides) on risk of incident dementia (all-cause dementia, Alzheimer's disease, and vascular dementia) using existing and new evidence.

@@

__Methods:__ Four distinct analyses were conducted. Firstly, a systematic review of 81 published and preprinted studies was used to summarise the existing evidence base. Preprints were searched using a new research tool created as part of this thesis. The existing evidence was then supplemented via two primary studies: i) a cohort study of the association of lipid regulating agents and dementia incidence in the Clinical Practice Research Datalink (CPRD); and ii) an individual participant data meta-analysis of the association of blood lipids with dementia incidence in previously unanalysed cohorts accessed through the Dementia Platform UK. Finally, a novel quantitative triangulation framework was proposed, building on recent developments in risk-of-bias assessment and bias-/indirectness-adjusted meta-analyses. This framework was then used to quantitatively triangulate the evidence identified and produced by this thesis.

@@

__Results:__ The systematic review did not identify a consistent effect of blood lipids on any dementia outcome across study designs, though there was some suggestion of a protective effect of LDL-c lowering on all-cause dementia and Alzheimer's disease in observational studies of statin use. The analysis of lipid regulating agents in the CPRD provided weak evidence for a protective effect of statins on all-cause dementia and Alzheimer's disease but suggested a harmful association with vascular dementia. However, the use of control outcomes illustrated this finding was likely due to confounding by indication related to vascular factors. The individual participant data meta-analysis suffered from a low response rate to data access requests, but in the three cohorts analysed, there was some evidence for only the association of triglycerides and vascular dementia. Finally, the triangulation analysis integrating the results of the previous three studies did not provide strong evidence for the causal effect of blood lipids on dementia outcomes.

@@

__Conclusions:__
This thesis provides new evidence concerning the role of blood lipids as a modifiable risk factor for dementia and highlights the uncertainty that still remains in relation to this causal question. In addition, it has developed new evidence synthesis methods and tools, specifically around the inclusion of preprints in systematic reviews and the quantitative triangulation of evidence sources.

>>>>

__Background:__ Identification of causal relationships between modifiable targets and risk of dementia is central to the development of evidence-based prevention strategies. As such, this thesis sought to triangulate the causal effects of lipid levels (total cholesterol, LDL-c, HDL-c, and triglycerides) on risk of incident dementia (all-cause dementia, Alzheimer's disease, and vascular dementia) using existing and new evidence.

@@

__Methods:__ Four distinct analyses were conducted. Firstly, a systematic review was used to summarise the evidence base. Existing evidence was then supplemented via two primary studies: i) a cohort study of lipid regulating agents and dementia incidence in the Clinical Practice Research Datalink (CPRD); and ii) an individual participant data (IPD) meta-analysis of lipid levels and dementia incidence. Finally, a novel quantitative triangulation framework was proposed, and was then used to integrate the results of the previous three studies.

@@

__Results:__ The systematic review did not identify a consistent effect of blood lipids on any dementia outcome across study designs considered, though there was some suggestion of a protective effect of LDL-c lowering on all-cause dementia and Alzheimer's disease in observational studies of statin use. Analysis of the CPRD data provided weak evidence for a protective effect of statins on all-cause dementia and Alzheimer's disease but suggested a harmful association with vascular dementia,  likely due to confounding by indication. The IPD meta-analysis provided some evidence for only the association of triglycerides and vascular dementia. Finally, the quantitative triangulation analysis did not provide strong evidence for the causal effect of blood lipids on dementia outcomes.

@@

__Conclusions:__
This thesis provides new evidence concerning the role of blood lipids as a modifiable risk factor for dementia and highlights the uncertainty that still remains in relation to this causal question. In addition, it has developed new evidence synthesis methods and tools.

### 2021-12-16 front-and-back-matter/_acknowledgements.Rmd
Other edits

There are many people without whose input and support this work would have been severely diminished:

Thank Julian separately -

First and foremost, I would like to thank my supervisors - Julian Higgins, Yoav Ben-Shlomo, Liz Coulthard and George Davey-Smith - for all their support over the past three years. Thanks also to Rupert Payne and Emma Anderson for their engaging discussion and useful feedback during my annual progress reviews. Sharen Hockey-O'Keffe helped me at numerous points throughout my PhD, and I was very lucky to receive her support. Similarly, Mark Newbury of the Dementia Platform UK worked tirelessly to support my applications for data access - I am extremely grateful for his help and commitment in chasing cohort owners.

Thanks must also go to those who contributed their time and energy to my research projects: Athena Sheppard, Emily Hennessy, Matthew Lee, Lena Schmidt, Alexandra McAleenan, Vincent Cheng, Francesca Spiga, Sarah Dawson and Rebecca Turner.

Thanks also to all those who raised issues or otherwise provided feedback on the open source tools built as part of this thesis:
<<github>> A special mention must go to Randall Boyes and Alex J Fowler for their input on the `robvis` tool.

@@

This thesis would not have been possible with Ciara Gardiner. Lots of people say this about someone at this point in the acknowledgements, and I'm always sceptical of whether it is true, so to avoid any doubt I mean it here quite literally.

Ciara, I'd need (at least) another <<words>> words to accurately describe all the ways in which you supported me through this endeavour. It can't have been fun being locked-down in a small flat with a grumpy and stressed PhD student, nor trying to support the same student remotely from Dublin. I talk about lessons learnt during this thesis in the Discussion, but the biggest takeaway for me is that I am incredibly lucky to have found you and to get to take on life while holding your hand. Thank you and I love you.

Thanks to Mum and Dad for your never-ending love and support, and everything they have done to get me to point where doing and submitting a thesis.

Thanks to Vivienne and Dad for giving up a substantial amount of time to proof-reading.

On a personal note, there are several people without whom I would not have survived the process of writing a dissertation during COVID times. Ben, Doug, Christin - thank you!

>>>>

First and foremost, I must thank my primary supervisor, Julian Higgins, for his tireless support and encouragement. I'd also like to thank my other supervisors - Yoav Ben-Shlomo, Liz Coulthard and George Davey-Smith - for all their input over the past three years. Thanks also to Rupert Payne and Emma Anderson for their engaging discussion and useful feedback during my annual progress reviews.

I am extremely grateful to all those who contributed their time and energy to my research projects: Athena Sheppard, Emily Hennessy, Matthew Lee, Lena Schmidt, Alexandra McAleenan, Vincent Cheng, Francesca Spiga, Sarah Dawson and Rebecca Turner. A special mention goes to Mark Newbury of the Dementia Platform UK, who worked tirelessly to support my applications for data access - I am extremely grateful for his help and commitment in chasing cohort owners.

I also want to thank the many people who raised issues or otherwise provided feedback on the open source tools built as part of this thesis - it's been a joy to watch others build on my work. A special mention must go to Randall Boyes and Alex J Fowler for their input/work on the `robvis` tool.

Ciara, I'd need (at least) another <<words>> words to accurately describe all the ways in which you supported me through this endeavour. It can't have been fun being locked-down in a small flat with a grumpy and stressed PhD student, nor trying to support the same student remotely from Dublin. In the Discussion, I talk about lessons learned during this thesis, but the biggest takeaway of this whole experience for me is that I am incredibly lucky to have found you and to get to take on life while holding your hand. Thank you and I love you.

Mum and Dad, thank you for your never-ending love and support. Everything I am comes from you, and you've done so much to get me to the point where a PhD thesis seemed possible.

There are several people without whom I would not have survived the process of writing a dissertation during COVID times. Ben, Doug, Christin - thank you!

Finally, a big thanks to Vivienne Gardiner, who sacrificed a substantial amount of time to proof-reading this thesis.

__Funding__

@@
